Pillar 1 - Sequential Pattern Mining

### 2019-02-12 Pillar 1 - Sequential Pattern Mining.Rmd
corrects make_sequences, creates SR

  sequences <- data %>%
    filter(
      !is.na(eventID) #this is here casue we have courses that were never taken.
      ) %>% 
    group_by(
      sequenceID,
      eventID
    ) %>%
    summarise(
      list_item = list(unique(!!enquo(item)))
    ) %>%
    ungroup %>%
    arrange(
      sequenceID,
      eventID
    ) %>%
    pull(
      list_item
      ) %>%
    as("transactions")

  # indicating sequence ID and event ID for each transaction
  sequences@itemsetInfo <- select(
    data,
    sequenceID,
    eventID
    ) %>%
    as.data.frame

@@

#
# Checking sequences
# inspect(head(sequences_G, 10))

rm(make_sequence, d_transactions)
```

@@

  rules %>%
    as("data.frame") %>%
    mutate(
      rule = str_remove_all(rule, pattern = "[<]"),
      rule = str_remove_all(rule, pattern = "[{]"),
      rule = str_remove_all(rule, pattern = "[}]"),
      rule = str_remove_all(rule, pattern = "[>]")
      ) %>%
    separate(rule, into = c("lhs", "rhs"), sep="=")  %>%
    select(
      lhs, rhs,
      support,
      confidence,
      lift
    ) %>%
    separate(lhs, into=c("LHS Course ID", "LHS Quality"), sep="_", remove= F) %>%
    separate(rhs, into=c("RHS Course ID", "RHS Quality"), sep="_", remove= F) %>%
    # Compute statistic
    group_by(lhs,`RHS Course ID`) %>%
    mutate(
      lhs.rhsCourse.support = sum(support),
      confidence = case_when(is.na(`RHS Quality`) ~ confidence,
                             T ~ (support / lhs.rhsCourse.support)),
      count = support * n_students,
      lhs.rhsCourse.count = lhs.rhsCourse.support * n_students,
      lift = 1 #dummy variable
      # TODO: compute rhs.support
      # lift = confidence.corr / rhs.support
     ) %>%
    ungroup() %>%
    mutate_at(
      c("support",
        "confidence", "lift"),
      funs(round(., 5))
    ) %>%
    filter(!`LHS Course ID`==`RHS Course ID`) %>%
    select(-`LHS Course ID`, -`LHS Quality`, -`RHS Course ID`, -`RHS Quality`)

@@

#Generating sequence rules:
SR_taken   <- my_SR(data = sequences$taken )
```

@@

##pass/fail filter
SR_PF      <- my_SR(data = sequences_PF) %>%
  filter(
    str_detect(rhs, "fail"),
    str_detect(lhs, "fail")
    )
```

@@

##high/low filter
SR_HL      <- my_SR(data = sequences_HL) %>% 
  filter(
    str_detect(rhs, "low"),
    str_detect(lhs, "low")
    )
```

@@

SR_Grade   <- my_SR(data = sequences_Grade )
```

@@

#Grades filter 
SR_expanded_Grade   <- my_SR(data = sequences_expanded_Grade )
```

SR_TPF are manually computed
```{r}
# current, past, not_yet
SR_TPF <- d_cumulative_transcript                  %>%
  select(sequenceID, eventID, item, course_not)    %>%
  unnest(item, .drop = F)                          %>%
  unnest(course_not)                               %>%
  left_join(
    select(d_transactions, 
           sequenceID, eventID, item, PF), 
    by= c("sequenceID", "eventID", "item")
    )                                              %>%
  select(sequenceID,eventID, course_not, item, PF) %>%
  rename(
    lhs="course_not", 
    rhs.x = "item", 
    rhs.PF = "PF")                                 %>%
  mutate(num_transactions = n())                   %>%
  group_by(lhs)                                    %>%
  mutate(
    num_lhs = n(),
    support_lhs         = num_lhs / num_transactions
    )                                              %>%
  group_by(rhs.x)                                  %>%
  mutate(
    num_rhs.x           = n(),
    support_rhs.x       = num_rhs.x / num_transactions
         )                                         %>%
  group_by(rhs.x,rhs.PF)                           %>%
  mutate(
    num_rhs.PF          = n(),
    support_rhs.PF      = num_rhs.PF / num_transactions
  )                                                %>%
  group_by(lhs, rhs.x)                             %>%
  mutate(
    num_rule_x          = n(),
    support_rule_rhs.x  = num_rule_x / num_transactions
  )                                                %>%
  group_by(lhs, rhs.x, rhs.PF)                     %>%
  mutate(
    num_rule_PF          = n(),
    support_rule_rhs.PF  = num_rule_PF / num_transactions,
    support_rhs_bounded  = num_rule_PF / num_rule_x
         )                                         %>%
  ungroup()                                        %>%
  mutate(
    confidence                   = support_rule_rhs.PF / support_lhs,
    confidence_bounded           = support_rhs_bounded / support_lhs, #not sure about this 
    lift                         = confidence / support_rhs.PF,
    lift_bounded                 = confidence_bounded /support_rhs_bounded #not sure about this
  )
```

We save and remove objects:
```{r saving and rm SR}
#Saving sequence rules
save(
    SR_taken, SR_PF, SR_HL, SR_Grade,
    SR_expanded_Grade,
    file = "App/SR.RDATA")

#clean unnecessary objects
rm(clean_SR, my_SR, 
   sequences, sequences_PF, sequences_HL, sequences_Grade,
   sequences_expanded_Grade,
   n_students,
   SR_taken, SR_PF, SR_HL, SR_Grade,

   SR_expanded_Grade)
```

>>>>

    # make transactions
  data_temp <- data %>%
    arrange(
      sequenceID,
      eventID
    ) %>%
    filter(
      !is.na(eventID) #this is here casue we have courses that were never taken.
      ) %>% 
    group_by(
      sequenceID,
      eventID
    ) %>%
    summarise(
      list_item = list(unique(!!enquo(item)))
    ) %>%
    ungroup

  sequences <- data_temp %>%
    pull(
      list_item
      ) %>%
    as("transactions")

  # indicating sequence ID and event ID for each transaction
  sequences@itemsetInfo <- data_temp %>%
    select(
      sequenceID,
      eventID
      ) %>%
    as.data.frame

@@

#
# Checking sequences
 inspect(head(sequences$G, 10))

rm(make_sequence #, d_transactions
   )
```

@@

  rules %>%
    as("data.frame") %>%
    mutate(
      rule = str_remove_all(rule, pattern = "[<]"),
      rule = str_remove_all(rule, pattern = "[{]"),
      rule = str_remove_all(rule, pattern = "[}]"),
      rule = str_remove_all(rule, pattern = "[>]")
      ) %>%
    separate(rule, into = c("lhs", "rhs"), sep=" = ")  %>%
    select(
      lhs, rhs,
      support,
      confidence,
      lift
    ) %>%
    separate(lhs, into=c("LHS Course ID", "LHS Quality"), sep="_", remove= F) %>%
    separate(rhs, into=c("RHS Course ID", "RHS Quality"), sep="_", remove= F) %>%
    # Compute statistic
    group_by(lhs,`RHS Course ID`) %>%
    mutate(
      lhs.rhsCourse.support = sum(support),
      confidence = case_when(is.na(`RHS Quality`) ~ confidence,
                             T ~ (support / lhs.rhsCourse.support)),
      count = support * n_students,
      lhs.rhsCourse.count = lhs.rhsCourse.support * n_students,
      lift = 1 #dummy variable
      # TODO: compute rhs.support
      # lift = confidence.corr / rhs.support
     ) %>%
    ungroup() %>%
    mutate_at(
      c("support",
        "confidence", "lift"),
      funs(round(., 5))
    ) %>%
    filter(!`LHS Course ID`==`RHS Course ID`) %>%
    select(-`LHS Course ID`, -`LHS Quality`, -`RHS Course ID`, -`RHS Quality`)

@@

#Generating sequence rules:
SR_taken   <- my_SR(data = sequences$taken)
```

@@

##pass/fail filter
SR_PF      <- my_SR(data = sequences$PF) %>%
  filter(
    str_detect(rhs, "fail"),
    str_detect(lhs, "fail")
    )
```

@@

##high/low filter
SR_HL      <- my_SR(data = sequences$HL) %>% 
  filter(
    str_detect(rhs, "low"),
    str_detect(lhs, "low")
    )
```

@@

SR_G   <- my_SR(data = sequences$G )
```

@@

#Grades filter 
#SR_expanded_Grade   <- my_SR(data = sequences_expanded_Grade )
```

SR_TPF are manually computed
```{r}
# current, past, not_yet
SR_TPF <- d_transcript_cum                         %>%
  select(sequenceID, eventID, course_current, course_not_so_far)    %>%
  unnest(course_current, .drop = F)                %>%
  unnest(course_not_so_far)                        %>%
  left_join(
    select(d_transactions, 
           sequenceID, eventID, item, PF), 
    by= c("sequenceID" = "sequenceID", "eventID" = "eventID", "item" = "course_so_far")
    )                                              %>%
  select(sequenceID,eventID, course_not_so_far, course_current, PF) %>%
  rename(
    lhs="course_not_so_far", 
    rhs.x = "course_current", 
    rhs.PF = "PF")                                 %>%
  mutate(num_transactions = n())                   %>%
  group_by(lhs)                                    %>%
  mutate(
    num_lhs = n(),
    support_lhs         = num_lhs / num_transactions
    )                                              %>%
  group_by(rhs.x)                                  %>%
  mutate(
    num_rhs.x           = n(),
    support_rhs.x       = num_rhs.x / num_transactions
         )                                         %>%
  group_by(rhs.x,rhs.PF)                           %>%
  mutate(
    num_rhs.PF          = n(),
    support_rhs.PF      = num_rhs.PF / num_transactions
  )                                                %>%
  group_by(lhs, rhs.x)                             %>%
  mutate(
    num_rule_x          = n(),
    support_rule_rhs.x  = num_rule_x / num_transactions
  )                                                %>%
  group_by(lhs, rhs.x, rhs.PF)                     %>%
  mutate(
    num_rule_PF          = n(),
    support_rule_rhs.PF  = num_rule_PF / num_transactions,
    support_rhs_bounded  = num_rule_PF / num_rule_x
         )                                         %>%
  ungroup()                                        %>%
  mutate(
    confidence                   = support_rule_rhs.PF / support_lhs,
    confidence_bounded           = support_rhs_bounded / support_lhs, #not sure about this 
    lift                         = confidence / support_rhs.PF,
    lift_bounded                 = confidence_bounded /support_rhs_bounded #not sure about this
  )
```

We save and remove objects:
```{r saving and rm SR}
#Saving sequence rules
save(
    SR_taken, SR_PF, SR_HL, SR_G,
    #SR_TPF,
    #SR_expanded_Grade,
    file = "App/SR.RDATA")

#clean unnecessary objects
# rm(clean_SR, my_SR, 
#    sequences,
#    #sequences_expanded_Grade,
#    n_students,
#    SR_taken, SR_PF, SR_HL, SR_G,
#    SR_TPF
#    #SR_expanded_Grade
#    )
```
